NGO analysis

### 2022-07-23 Shiny/Test/Flexdashboard rows.Rmd
File input validation. Adding KPI equations.

The KPIs are based on the following calculations:

- **admin** is based on dividing - expenses_administrative / expenses_program_services *100
- **Fundraising ROI** - (revenue_private_contributions +
                 revenue_membership_fees - expenses_fundraising) /
              expenses_fundraising * 100
```{r}

@@

### Plots
```{r}

@@

  #---- Data frames ------------------------------------------------------------      
  # creating data frame from html file
  htmldata <- reactive({
    read_proff(
      input$html$datapath,
      as.character(input$start), as.character(input$end)
    )
  })

  # Creating data frame from Excel file
  exceldata <- reactive({
    widen_excel(input$excel$datapath)
  })

@@

  #----Revenue------------------------------------------
  revenue_plot <- reactive({
    # testing for data quality
    if (datatest() == TRUE) {
      # printing 'bad data' plot if data are bad
      bad_data()
    } else {
      exceldata() |>
        # Renaming for the purposes of the plot + dropping other variables
        transmute(
          Year = as.factor(year),
          `Private contributions` = revenue_private_contributions,
          `Membership fees` = revenue_membership_fees,
          `Public grants` = revenue_public_grants,
          Investments = revenue_investments,
          `Operational activities` = revenue_operational_activities,
          `Commercial activities` = revenue_commercial_activities,
          Other = revenue_other
        ) |>
        # Pivoting to use fill
        pivot_longer(
          c(`Private contributions`:Other),
          names_to = "Revenue source"
        ) |>
        mutate(
          `Revenue source` =
            ordered(`Revenue source`, levels =
                      c("Other", "Commercial activities",
                        "Operational activities", "Investments",
                        "Membership fees", "Private contributions",
                        "Public grants"))
        ) |> 
        ggplot(aes(Year, value, fill = `Revenue source`)) +
        geom_col() +
        scale_y_continuous(
          # Excel data are assumed to be in NOK; if this changes, code based on
          # the currency_code() function may be adapted
          labels = label_dollar(prefix = "NOK "),
          # Basing lower limit on the data (not suppressing erroneous negative
          # numbers), and the upper limit on the maximum value in revenues and
          # expenses, to match the two plots
          limits = c(NA, maximum())
        ) +
        # Specifying the number of values, as the base palette only includes 6
        scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
        labs(x = "", y = "", title = "Revenue (in thousands)")
    }
  })

>>>>

The KPIs are based on the following calculations, using data derived from the Excel sheet:

- **Fundraising ROI**: $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$
- **Liquid assets/obligations**: $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
- **Private donations growth**: $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$
- **Public grants growth**: $$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
- **Private/public proportion**: $$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$

@@

### Plots

@@

  #---- Data frames ------------------------------------------------------------      
  # creating data frame from html file
  htmldata <- reactive({
    validate(need(tools::file_ext(input$html$datapath) == "html",
                  "Please upload an html file."))
    read_proff(
      input$html$datapath,
      as.character(input$start), as.character(input$end)
    )
  })

  # Creating data frame from Excel file
  exceldata <- reactive({
    validate(need(tools::file_ext(input$excel$datapath) == "xlsx",
                  "Please upload an Excel file."))
    widen_excel(input$excel$datapath)
  })

@@

  #----Revenue------------------------------------------
  revenue_plot <- reactive({
    # testing for data quality
    if (datatest() == TRUE) {
      # printing 'bad data' plot if data are bad
      bad_data()
    } else {
      exceldata() |>
        # Renaming for the purposes of the plot + dropping other variables
        transmute(
          Year = as.factor(year),
          `Private contributions` = revenue_private_contributions,
          `Membership fees` = revenue_membership_fees,
          `Public grants` = revenue_public_grants,
          Investments = revenue_investments,
          `Operational activities` = revenue_operational_activities,
          `Commercial activities` = revenue_commercial_activities,
          Other = revenue_other
        ) |>
        # Pivoting to use fill
        pivot_longer(
          c(`Private contributions`:Other),
          names_to = "Revenue source"
        ) |>
        # Ordering for the plot
        mutate(
          `Revenue source` =
            ordered(`Revenue source`, levels =
                      c("Other", "Commercial activities",
                        "Operational activities", "Investments",
                        "Membership fees", "Private contributions",
                        "Public grants"))
        ) |> 
        ggplot(aes(Year, value, fill = `Revenue source`)) +
        geom_col() +
        scale_y_continuous(
          # Excel data are assumed to be in NOK; if this changes, code based on
          # the currency_code() function may be adapted
          labels = label_dollar(prefix = "NOK "),
          # Basing lower limit on the data (not suppressing erroneous negative
          # numbers), and the upper limit on the maximum value in revenues and
          # expenses, to match the two plots
          limits = c(NA, maximum())
        ) +
        # Specifying the number of values, as the base palette only includes 6
        scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
        labs(x = "", y = "", title = "Revenue (in thousands)")
    }
  })
